---
title: "Diamonds dataset Analysis"
author: "Murtaza Badshah"
output: html_document
---

### Loading packages

```{r}
library(MASS)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```
\newpage
### Problem 1. For this problem we'll use the `diamonds` dataset from the `ggplot2` package. Use the functions inside `dplyr` and `ggplot2` packages to answer the following questions. 

```{r}
# Change the object of diamonds dataset from tibble to data frame.
diamonds <- as.data.frame(diamonds)
```

##### (a) Draw a bar chart to display the total number of diamonds in the `diamonds` dataset, grouped by `color` grading.
```{r}
diamonds%>%
  ggplot(aes(color))+
  geom_bar(aes(group = color, fill = color))+
  coord_flip()+
  geom_text(stat='count', aes(label=..count..), vjust=0.2, hjust = 1.5, 
            colour = "darkred", fontface = "bold")
```

##### (b) We call diamond is colorless if color grading is D, E, or F. Select colorless diamonds from the `diamonds` dataset and assign it to `colorless_diamonds` for later use.
```{r}
colorless_diamonds <- diamonds%>%
                          filter(color %in% c("D","E","F"))
```
\newpage
##### (c) Use `colorless_diamonds` data frame to recreate the uploaded *"The Value of Colorless Diamond"* plot.
```{r}
  colorless_diamonds%>%
  ggplot(aes(carat,price, color = clarity))+
  geom_jitter()+
  geom_smooth()+
  scale_color_brewer(palette = "Spectral")+
  labs(x="Weight of the diamond (carat)", y = "Price (US Dollars)", 
       title = "The Value of Colorless Diamond")
```

##### (d) The ideal depth for round diamond is between two percentages 59.5\% and 62.9\%, inclusive. Select the ideal depth diamonds from the `diamonds` dataset and assign it to `ideal_depth_diamonds` for later use.
```{r}
ideal_dept_diamonds <- diamonds%>%
  filter(diamonds$depth >= 59.5 & diamonds$depth <= 62.9)
```
\newpage
##### (e) Use `ideal_depth_diamonds` data frame to draw the histogram of the `depth` for each quality of the `cut` separately in one graphic window. 
```{r}
ideal_dept_diamonds%>%
  ggplot(aes(depth))+
  geom_histogram(bins = 30)+
  labs(x = "Depth of Diamonds", 
       title = "Histogram of diamond cuts based on depth.")+
  facet_wrap(~ cut)
```

##### (f) The length and width ratio gives an idea how big or small the diamond will appear, also defining its shape. To calculate the length and width ratio just divide the length (`x`) of the diamond by its width (`y`). Add a new variable named `LWratio` to the `diamonds` dataset and assign it to `LW_diamonds`.
```{r}
LW_diamonds <- diamonds%>%
  mutate(LWratio = x/y)
```

